### 2022-09-17 main/chapter_searches.tex
remove significance section; this is already far too long; some tweaks

How can one escape this propaganda?
Likelihood ratios on the actual data escape immediately by bypassing the
censor.
Even inequalities need not be biased --- if taken against a predetermined
threshold, say $5$ events, then models can predict whether $n \leq 5$,
and its truth or falsehood in data can favour either side.%
\footnote{%
This context of a fixed inequality is one condition where the
\pvalue\ ratio is not misleading when interpreted as a likelihood ratio.%
}
Finally, a reader can fall back to not trusting the reporter ---
the statements `\atlas\ reports $n \leq 5$' and logical `$n \leq 5$' are not
identical, since one can infer from the former that in fact $n = 5$.
One knows what the message says, but not that it states the whole truth.

@@

\subsection{Gaussian significance}
Standard-deviations distance is a more intuitive measure of compatibility than
a \pvalue.

defining $\chi = (x - \mu) / \sigma$, the Gaussian likelihood is

\begin{equation}
L(\chi) =
\frac{1}{\sqrt{2\pi}}
\exp\!\left(-\frac{1}{2}\chi^2\right)
\end{equation}

\TODO{log maximum likelihood ratio}
$\log$ maximum likelihood ratio $\kappa$
\begin{equation}
\kappa = -\frac{1}{2}\chi^2
\quad
\Leftrightarrow
\quad
\chi = \pm\sqrt{-2\kappa}
\end{equation}

\TODO{tail area}

\begin{align}
P(\chi)
&=
\frac{1}{\sqrt{2\pi}}
\int_\chi^\infty
\!
\exp\!\left(-\frac{1}{2}t^2\right)
\,\mathrm{d}t
\\[0.5em]
&=
L(\chi)
\times
\left(
\frac{1}{\chi}
+
\mathcal{O}\!\left(\frac{1}{\chi^3}\right)
\right)
\end{align}
\TODO{significance measures}
~\cite{
jaynes2003probability,
abramowitz1965handbook
}

% significance measures
ATLAS' recommended significance measure~\cite{atlas_significance} is
\begin{align}
\label{eqn:significance_atlas}
S_\mathrm{\atlas}(n; \mu, \sigma) =~&
\sqrt{2} \times
\mathrm{sign}(n - \mu) \times
\\[0.2em] \nonumber
&
\sqrt{
n\log\left(\frac{n(\mu + \sigma^2)}{\mu^2 + n\sigma^2}\right)
- \frac{\mu^2}{\sigma^2}\log\left(
1 + \frac{\sigma^2(n - \mu)}{\mu(\mu + \sigma^2)}
\right)
}
,
\end{align}
taken in a limit where $0\log(0x) = 0$ for the $n=0$ case.

% similarity and difference?? TODO

>>>>

How can one escape this trap?
Likelihood ratios on the actual data escape immediately by bypassing the
censor, but even inequalities need not be biased --- if taken against a
predetermined threshold, say $5$ events, then models can predict whether
$n \leq 5$, and the truth or falsehood in data can favour either side.%
\footnote{%
This context of a fixed inequality is one condition where the
\pvalue\ ratio is not misleading when interpreted as a likelihood ratio.%
}
Ultimately, the reader can fall back to not trusting the reporter ---
the statements `\atlas\ reports $n \leq 5$' and `logically, $n \leq 5$' are not
identical, since one might infer from the former that in fact $n = 5$.
The message itself is data --- what it says may not be the whole truth.

@@

### 2022-09-17 main/draft.tex
remove significance section; this is already far too long; some tweaks

>>>>

% whooo removing a section

\subsection{Gaussian significance}
Gaussian standard-deviations distance is an intuitive measure of compatibility
with a prediction.
than
a \pvalue.

defining $\chi = (x - \mu) / \sigma$, the Gaussian likelihood is

\begin{equation}
L(\chi) =
\frac{1}{\sqrt{2\pi}}
\exp\!\left(-\frac{1}{2}\chi^2\right)
\end{equation}

\TODO{log maximum likelihood ratio}
$\log$ maximum likelihood ratio $\kappa$
\begin{equation}
\kappa = -\frac{1}{2}\chi^2
\quad
\Leftrightarrow
\quad
\chi = \pm\sqrt{-2\kappa}
\end{equation}

\TODO{tail area}

\begin{align}
P(\chi)
&=
\frac{1}{\sqrt{2\pi}}
\int_\chi^\infty
\!
\exp\!\left(-\frac{1}{2}t^2\right)
\,\mathrm{d}t
\\[0.5em]
&=
L(\chi)
\times
\left(
\frac{1}{\chi}
+
\mathcal{O}\!\left(\frac{1}{\chi^3}\right)
\right)
\end{align}
\TODO{significance measures}
~\cite{
jaynes2003probability,
abramowitz1965handbook
}

% significance measures
ATLAS' recommended significance measure~\cite{atlas_significance} is
\begin{align}
\label{eqn:significance_atlas}
S_\mathrm{\atlas}(n; \mu, \sigma) =~&
\sqrt{2} \times
\mathrm{sign}(n - \mu) \times
\\[0.2em] \nonumber
&
\sqrt{
n\log\left(\frac{n(\mu + \sigma^2)}{\mu^2 + n\sigma^2}\right)
- \frac{\mu^2}{\sigma^2}\log\left(
1 + \frac{\sigma^2(n - \mu)}{\mu(\mu + \sigma^2)}
\right)
}
,
\end{align}
taken in a limit where $0\log(0x) = 0$ for the $n=0$ case.

% similarity and difference?? TODO
